---
title: "`dplyr` 2: Pipes and Groups"
---

```{r}
library(tidyverse)
library(ggthemes)
library(nycflights23)
```

# Wrangling the `nycflights23` data

## `summarize()`

Using the `flights` data from the {nycflights23} package, use `summarize()` to compute statistics about the data:

1.  The lowest and highest `distance` traveled

2.  The lowest and highest `air_time`

3.  The median `dep_delay`

```{r}
flights
```

## `summarize()` II

Using the `flights` data from the {nycflights23} package, extract the rows for Delta flights. (Hint: the carrier code is “DL”)

Then use `summarize()` and a summary function to find:

1.  The number of flights in this subset

2.  The median `dep_delay`. How does it compare to the overall median?


```{r}
flights
```

## `group_by()`

Use `group_by()`, `summarize()`, and `slice_max()` to display the five `dest` airports with the most flights from NYC airports in 2023. Your display should include the median flight delay for these airports.

```{r}
flights
```

# Intro to Tidy Data

```{r}
bakeoff_ratings <- read_csv("https://stat220-w26.github.io/data/bakeoff_messy_ratings.csv")
bakeoff_ratings
```

## Discuss

Is this dataset in tidy format? Why or why not?

If not, what would a tidy data set look like? Sketch out the first few rows of this data set in tidy format

## `tidy`ing bakeoff ratings

Follow along with the slides to tidy the `bakeoff_ratings` dataset

# `messyratings2`

```{r}
messy_ratings2 <- read_csv("https://stat220-w26.github.io/data/messy_ratings2.csv")
messy_ratings2
```

## Task 1

Tidy this data set by

1.  Selecting the `series` and `e*_7day` columns
2.  Pivoting the data to add a column for `episode` and a column for `rating` (we'll clean up the episode column later)

```{r}
# include your code here
```

## Task 2

-   Clean the `episode` and `period` column
-   Make a *line plot* with episode on the x-axis, rating on the y-axis, colored by series. (You will also need to map the `group` aesthetic to series)

