---
title: "`tidyr` 2: Reshaping Data"
---

```{r}
library(tidyverse)
library(ggthemes)
```


# Last time

Here is a data frame with the ratings for each episode of The Great British Bakeoff for series 1-8.

```{r}
bakeoff_ratings <- read_csv("https://stat220-w26.github.io/data/bakeoff_messy_ratings.csv")
bakeoff_ratings
```

and here is the code we used last time to tidy the data: 

```{r}
bakeoff_ratings %>%
  pivot_longer(cols = e1:e10, names_to = "episode", values_to = "rating") %>%
  mutate(
    episode = parse_number(episode) 
  )
```

# `messyratings2`

```{r}
messy_ratings2 <- read_csv("https://stat220-w26.github.io/data/messy_ratings2.csv")
messy_ratings2
```

## Task 1

Tidy this data set by

1.  Selecting the `series` and `e*_7day` columns
2.  Pivoting the data to add a column for `episode` and a column for `rating` (we'll clean up the episode column later)

```{r}
# include your code here
```

## Task 2

::: {.task .nonincremental}
-   Clean the `episode` and `period` column
-   Make a *line plot* with episode on the x-axis, rating on the y-axis, colored by series. (You will also need to map the `group` aesthetic to series)
:::

# More Practice

## `relig_income`

The `relig_income` dataset in the {tidyr} package stores counts based on a survey which (among other things) asked people about their religion and annual income:

```{r}
relig_income
```

Use `pivot_longer()` to tidy this dataset.

## `anscombe`

[Anscombe's quartet](https://en.wikipedia.org/wiki/Anscombe%27s_quartet) is a built-in dataset in R.

```{r}
anscombe
```

This dataset contains four pairs of variables (`x1` and `y1`, `x2` and `y2`, etc) that underlie Anscombe’s quartet, a collection of four datasets that have the same summary statistics (`mean`, `sd`, `correlation` etc), but have quite different data. We want to produce a dataset with columns `set`, `x` and `y`:

```         
# A tibble: 44 × 3
   set       x     y
   <chr> <dbl> <dbl>
 1 1        10  8.04
 2 2        10  9.14
 3 3        10  7.46
```

There are (at least) two ways to do this. The first is a little more intuitive, but not as efficient:

1.  First, we'll create a new "index" column, so we don't lose track of which x values map to which y values.

```{r}
anscombe <- anscombe %>%
  mutate(
    index = 1:nrow(anscombe)
  )
```

2.  Next, use `pivot_longer()` on all of the columns but `index`. Use the default for `names_to` and `values_to`. The first few rows of your result should look like this:

```         
# A tibble: 88 × 3
   index name  value
   <int> <chr> <dbl>
 1     1 x1    10   
 2     1 x2    10   
 3     1 x3    10  
```

3.  Next, separate `name` into `variable` and `set`. This is a little tricky, since there's no separator character (the values of `name` are `x1` and `x2` instead of `x_1` or `x_2`). Instead, set `sep = 1`, which tells R to split the column after the first character. The first few rows of your result should look like this:

```         
# A tibble: 88 × 4
   index variable set   value
   <int> <chr>    <chr> <dbl>
 1     1 x        1     10   
 2     1 x        2     10   
 3     1 x        3     10   
```

4.  Finally, use `pivot_wider()` with `names_from` `variable` and `values_from` `value`. Call your tidy dataset `anscombe_tidy`.

If all went well, you should be able to run the following two chunks to generate the summary statistics and scatterplots for Anscombe's quartet.

```{r}
#| eval: false


anscombe_tidy %>%
  group_by(set) %>%
  summarize(
    mean_x = mean(x),
    mean_y = mean(y),
    sd_x = sd(x),
    sd_y = sd(y),
    cor = cor(x,y)
  )
```

```{r}
#| eval: false

anscombe_tidy %>%
  ggplot(aes(x = x, y = y)) + 
  geom_point() + 
  facet_wrap(~set)
```

The second way to do this is directly within `pivot_longer()`:

```{r}
anscombe %>%
  pivot_longer(
    -index,
    names_to = c(".value", "set"),
    names_pattern = "(.)(.)"
  )
```

pull up the help page for `pivot_longer` and try to explain the new arguments.
