---
title: "`bind_` and `_join`: Combining Data"
---

```{r}
library(tidyverse)
library(ggthemes)
```

Click the "code" button above to copy and paste the source code.

# Warm Up

```{r}
load(url("https://stat220kurtz.github.io/data/combining-data-examples.Rda"))
```

## Example 1: Star Wars Characters 



Dataset 1: 

- Each row in the dataset corresponds to a Star Wars character
- The columns include their `name`, `height`, `mass`, `homeworld`, `species`, and the `first_film` they appeared in

```{r}
#| echo: true
starwars_characters
```

Dataset 2:

The dataset is formatted the same, but includes two characters that were first introduced in *The Last Jedi*, a Star Wars movie that came out after the first dataset was released. 

```{r}
#| echo: true
starwars_lastjedi
```

How should these datasets be combined?

## Example 2: Stats Sections

Dataset 1:

This dataset includes the number of sections of stats courses offered in Fall 2024 and Winter 2025 terms.

```{r}
#| echo: true
stats_sections_fw
```

Dataset 2: 

This dataset includes the number of sections of stats courses offered in the Spring 2025 term.

```{r}
#| echo: true
stats_sections_s
```

How should these datasets be combined?


## Example 3: Survivor castaways

Dataset 1:

- This dataset includes information about each castaway's performance in a given season of *Survivor* (US seasons)
- It includes a castaway ID, first name, season name and number, and the results for that castaway in that season (their overall `place`, whether they made the `jury`, and whether they were a `finalist`)

```{r}
#| echo: true
us_castaway_results
```

Dataset 2:

- This dataset includes information about each castaway who has appeared on *Survivor* (including non-US seasons)
- It includes a castaway ID, their full name, date of birth, gender and occupation as presented in the show, and their Myers-Briggs personality type.  

```{r}
#| echo: true
cast_details
```

We are interested in exploring the relationship between overall place on a Survivor season and personality type. How should these datasets be combined?



# Bakeoff Ratings (again)

We'll start with our `messy_ratings2` dataset from last week:

```{r}
messy_ratings2 <- read_csv("https://stat220kurtz.github.io/data/messy_ratings2.csv")
messy_ratings2
```

On Friday, we tidied the data[^1] using the following code:

[^1]: Talk it over: does everybody remember what makes a dataset "tidy"?

-   `select` columns of interest
-   `pivot_longer` to extract the variable encoded in the column names
-   `separate` to separate episode number from ratings period
-   `parse_number` to format columns as numeric

```{r}
tidy_ratings2 <- messy_ratings2 %>%
  select(series, contains("7day")) %>%
  pivot_longer(-series, names_to = "episode", values_to = "viewers") %>%
  separate(episode, into = c("episode", "period")) %>%
  mutate(
    episode = parse_number(episode),
    period = parse_number(period)
  )

tidy_ratings2
```


## Task: Add series 9-14 to the dataset

The code chunk below loads the new dataset

```{r}
messy_ratings2b <- read_csv("https://stat220kurtz.github.io/data/messy_ratings2_9_14.csv")
messy_ratings2b
```

### Combine with `bind_rows()`

Combine the two **messy** datasets by using `bind_rows()`. Be sure to look at your resulting dataset and make sure it has the right format

```{r}
# your code here
```

### Combine with `join()`

We'll practice `join`ing with the **tidy** versions of the data. First, create `tidy_ratings2b` using the new dataset. Follow the same steps as what was used to create `tidy_ratings2`

```{r}
# your code here
```

Once you're satisfied that the two tidy datasets look similar enough to combine, remove `#| eval: false` from the following code chunk and run it. Did it work? Why or why not?

```{r}
#| eval: false
tidy_ratings = tidy_ratings2 %>%
  left_join(tidy_ratings2b)
```

Use the correct `_join` function here. Save your dataset to the `tidy_ratings` object.

```{r}
# your code here
```

### Combine with `bind_cols()`

To join the two datasets using `bind_cols()`, you'll first have to un-tidy your tidy datasets, so that each episode number is in a *row* and each series number is in a *column*. Call the datasets you create by doing this `untidy_ratings2` and `untidy_ratings2b`. After using `bind_cols()` the first few rows of your final dataset should look like this:

```         
# A tibble: 10 × 16
   episode period    s1    s2    s3    s4    s5    s6    s7    s8    s9   s10   s11   s12   s13   s14
     <dbl>  <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl> <dbl>
 1       1      7  2.24  3.1   3.85  6.6   8.51  11.6  13.6  9.46  9.55  9.62  11.2  9.57  8.3   7.84
 2       2      7  3     3.53  4.6   6.65  8.79  11.6  13.4  9.23  9.31  9.38  10.8  8.98  7.6   7.54
 3       3      7  3     3.82  4.53  7.17  9.28  12.0  13.0  8.68  8.91  8.94  10.7  8.69  7.35  7.28
```

**Two hints:** 1) Use `pivot_wider()` to untidy the datasets - the names_from and values_from parameters will need to be specified. 2) After creating the untidy datasets but before merging them using `bind_cols()`, remove the `episode` and `period` columns from `untidy_ratings2b`. Why did I suggest you do this?

```{r}
# your code here
```

# `join` the episodes data

The `episodes` data has information about each episode of GBBO. In particular, it includes:

-   `baker`: each baker that competed on that episode
-   `signature`: the name of the dessert they made for the signature challenge
-   `technical`: the place the baker earned in the technical challenge
-   `showstopper`: the name of the dessert they made for the showstopper challenge
-   `result`: whether the baker was "Safe" or "Eliminated"

```{r}
episodes = read_csv("https://stat220kurtz.github.io/data/bakeoff-episodes.csv")
episodes
```

Join the tidy dataset (all 14 seasons) with the `episodes` data.

Your task is to print the "signature" desserts that were made on the 10 episodes that had the highest 7-day viewership.

```{r}
# your code here
```
