---
title: "Data import and dates/times"
---

```{r}
library(tidyverse)
library(ggthemes)
library(readxl)
```

Click the "code" button above to copy and paste the source code into RStudio

# Warm Up

Use `read_csv()` to import the `desserts` data set from

https://stat220kurtz.github.io/data/desserts.csv

```{r}
# your code here
```


# `readr` practice

Use the appropriate `read_<type>()` function to import the following data sets:

- `https://stat220kurtz.github.io/data/data-4.csv`

- `https://stat220kurtz.github.io/data/tricky-1.csv`

If you hit any errors/problems, be sure to explore them and identify the issue, even if you can't "fix" it.

## data-4.csv

```{r}
# your code here
```

## tricky-1.csv

```{r}
# your code here
```


# `read_sheet` practice

Use the `read_sheet` function from {googlesheets4} to import the following data set:

- `https://docs.google.com/spreadsheets/d/1NN2S7bHSQbiHgd6gJnMUoJ6_eB-YgqLi_JXckFzTkbc/edit?usp=sharing`

Step 1: read in the data so it looks like the following:

```
# A tibble: 9 × 2
  id      n    
  <chr>   <chr>
1 Brand 1 n    
2 1234    8    
3 8721    2    
4 1822    3    
5 Brand 2 n    
6 3333    1    
7 2156    3    
8 3987    6    
9 3216    5   
```

```{r}
library(googlesheets4)
# your code here
```

Step 2 (Stretch goal): Manipulate the data so that it looks like the following:

```
# A tibble: 7 × 3
  brand   id    n    
  <chr>   <chr> <chr>
1 Brand 1 1234  8    
2 Brand 1 8721  2    
3 Brand 1 1822  3    
4 Brand 2 3333  1    
5 Brand 2 2156  3    
6 Brand 2 3987  6    
7 Brand 2 3216  5  
```

```{r}
# your code here
```


::: callout-tip
## Hints

(1) We need to create a `brand` column. Start by creating this column, filling in the value from the "id" column if `n == "n"` and NA otherwise
(2) Next, we want to "fill in" all of the NA's in our "brand" column. Look at the help page for the `fill` function
(3) Use `filter` to remove the redundant rows and `select` to remove the redundant columns
:::

# `lubridate` practice

## Task 1

Create a new copy of the `desserts` dataset, but do *not* parse the `uk_airdate` within `read_csv`. Instead, leave it as a character vector and parse the date using {lubridate} functions. Which approach do you prefer?

Then, create a new column called `how_long_ago` that measures the time between today and the UK airdate of the episode. Can you format this column:

  - in years
  - in months
  - in weeks
  - in days
  
(*Hint:* see `?time_length`)

```{r}
# your code here
```

