---
title: "12-factors"
---

```{r}

library(tidyverse)
library(lubridate)
library(patchwork)
library(ggthemes)
```

# Warm Up

`gss_cat` gets loaded with the `forcats` package in the `tidyverse`. Use `gss_cat` to answer the following questions (I'd use a graph, but do whatever you'd like!) 

```{r}
gss_cat
```

1. Which religions watch the least TV?

2. Do married people watch more or less TV than single people?


# Try it: `fct_reorder`

Use `rincome_summary` to construct a dotplot of `rincome` against `age`.

Reorder `rincome` by `age`

```{r}
rincome_summary <- gss_cat %>%
  group_by(rincome) %>%
  summarize(
    age = mean(age, na.rm = TRUE),
    tvhours = mean(tvhours, na.rm = TRUE),
    n = n()
  )
```


# Try it `fct_collapse`

Collapse the `marital` variable to have levels `Married`, `not_married`, and `No answer`

Include `"Never married"`, `"Divorced"`, and "`Widowed"` in `not_married`

# Additional examples 

## Hotel bookings by month

```{r}
# From TidyTuesday: https://github.com/rfordatascience/tidytuesday/blob/master/data/2020/2020-02-11/readme.md
hotels <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-02-11/hotels.csv", show_col_types = FALSE)
```

1. Run the code and examine the plot. How are the months ordered? What would be a better order?

```{r fig.width=10}
hotels %>%
  group_by(hotel, arrival_date_month) %>%   # group by hotel type and arrival month
  summarize(mean_adr = mean(adr)) %>%       # calculate mean adr for each group
  ggplot(aes(
    x = arrival_date_month,                 # x-axis = arrival_date_month
    y = mean_adr,                           # y-axis = mean_adr calculated above
    group = hotel,                          # group lines by hotel type
    color = hotel)                          # and color by hotel type
    ) +
  geom_line() +                             # use lines to represent data
  theme_minimal() +                         # use a minimal theme
  labs(
    x = "Arrival month",                    # customize labels
    y = "Mean ADR (average daily rate)",
    title = "Comparison of resort and city hotel prices across months",
    subtitle = "Resort hotel prices soar in the summer while ciry hotel prices remain relatively constant throughout the year",
    color = "Hotel type"
    )
```


2. Reorder the months on the x-axis (levels of `arrival_date_month`) in a way that makes more sense. Use a function from the `forcats` package.

## More hotel bookings by room type

1. Create a bar chart of the `reserved_room_type`.

2. Create another bar chart where you have bars for the top three room types and an "other" category lumping the other room types together.


## More hotel bookings by room type

1. Run the code and examine the plot. How are the months ordered? What would be a better order?


```{r fig.width=10}
hotels %>%
  group_by(meal, arrival_date_month) %>%   # group by meal type and arrival month
  summarize(mean_adr = mean(adr)) %>%       # calculate mean adr for each group
  ggplot(aes(
    x = arrival_date_month,                 # x-axis = arrival_date_month
    y = mean_adr,                           # y-axis = mean_adr calculated above
    group = meal,                          # group lines by meal type
    color = meal)                          # and color by meal type
    ) +
  geom_line() +                             # use lines to represent data
  theme_minimal() +                         # use a minimal theme
  labs(
    x = "Arrival month",                    # customize labels
    y = "Mean ADR (average daily rate)",
    color = "Meal type"
    ) +
  ggthemes::scale_color_colorblind()
```

2. Reorder the levels of meal type in a way that makes more sense. Use a function from the `forcats` package to do this.

