---
title: "13: Intro to Strings/Regex"
---

```{r}
library(babynames)
library(tidyverse)
```

# `babynames` intro

Replicate the "Amanda" plot with your own name using the starter code below. If your name doesn't have enough data, try a friend or professor's name

```{r}
#| eval: false 
babynames %>%
  ___(name == ___ ) %>%
  ggplot(aes(x = ___, y = ___, col = ___)) +
  geom____() 
```

# Practice with `stringr`

## Your turn 

No code needed, just think about what it returns


## Your turn 2

Fill in the blanks of the .Rmd file to...

1. Isolate the last letter of every name

2. and create a logical variable that displays whether the last letter is one of "a", "e", "i", "o", "u", or "y".

3. Use a weighted mean to calculate the proportion of children whose name ends in a vowel, by year (see `?weighted.mean`).

4. and then display the results as a line plot.

```{r}
#| eval: false # Remember to remove this line


babynames %>%
  mutate(last = ___, 
         vowel = ___) %>%
  group_by(___) %>%
  ___(p_vowel = weighted.mean(vowel, n)) %>%
  ___ +
  ___
```


## Example: Carleton courses

The below code chunk imports a data set of the 6-credit courses offered at Carleton in Winter 2023. All three columns are character vectors.

```{r}
courses <- read_csv("https://stat220-w26.github.io/data/winter2023_course_tbl.csv", col_types = list(course = col_character()))
```

### (a)

How many of the course numbers end in `.00`? Use `str_detect()` or `str_count()` to help you answer this question.

Note that `.` is a special character in strings, so use `\\.` to get the literal period.


### (b)

The section number appears after the decimal point. Use `mutate()` and `str_sub()` to create a `section` column containing this number.


### (c)

How many courses contain the word `Introduction`? Does case matter here?


### (d)

What is the longest course name (in terms of characters)? What is the shortest course name? Use `str_length()` to help you answer this question.

### (e)

Use `str_subset()` to return the course names that contain exclamation points (`!`).

# Practice with Regular Expressions

## Your turn 1

Detect  either "." or "-" in the `info` vector.

```{r}
a1 <- "Home: 507-645-5489"
a2 <- "Cell: 219.917.9871"
a3 <- "My work phone is 507-202-2332"
a4 <- "I don't have a phone"
info <- c(a1, a2, a3, a4)
```


## Your turn: ends with a vowel

Fill in the code to determine how many baby names in 2015 ended with a vowel.

```{r}
#| eval: false # remember to remove this line


babynames %>% 
  ___(___ == ___) %>%                       # extract year 2015
  ___(ends_with_vowel = ___(___, ___)) %>%  # create logical column
  count(ends_with_vowel)                    # create a frequency table
```

## Additional practice (if time)

A vector called `words` is loaded with `stringr` and contains a corpus of 980 words used in text analysis. Use `stringr` functions and regular expressions to find the words that satisfy the following descriptions. 


#### 1. Find all words that start with `y`.

```{r}
pattern <- "type your pattern here"
str_subset(words, pattern)
```

#### 2. Find all words that end with `x`.

#### 3. Find all words that are exactly three letters long.

#### 4. Find all words that start with a vowel.

#### 5. Find all words that start with consonants.

#### 7. Find all words that end with `ing` or `ise`.

#### 8. Find all words that have seven letters or more.

#### 9. Find all words that start with three consonants.

