---
title: "14-text-case-study"
---

```{r}
library(tidyverse)
library(tidytext) # functions for doing text analysis
```

# Load the Data

```{r}
en_coursera_reviews <- read_csv("https://stat220-w26.github.io/data/en_coursera_sample.csv")
glimpse(en_coursera_reviews)
```


# Tokenize

```{r}
library(tidytext)
en_coursera_reviews |>
  unnest_tokens(output = word, input = Review) |>
  select(CourseId, word)
```

# Count the tokens

```{r}
en_coursera_reviews |>
  unnest_tokens(output = word, input = Review) |>
  anti_join(get_stopwords(source = "stopwords-iso"), by = "word") |>
  filter(str_detect(CourseId, "data-science")) |> 
  group_by(CourseId) %>%
  count(word) %>%
  slice_max(n, n = 10) %>%
  ggplot(aes(x = n, y = fct_reorder(word, n))) + 
  geom_col() + 
  facet_wrap(~CourseId, scales = "free_y", nrow = 2) + 
  labs(
    y = "Word"
  )
```

# Create "clean" reviews

```{r}
coursera_reviews <- en_coursera_reviews |>
  unnest_tokens(word, Review) |>
  anti_join(get_stopwords(source = "stopwords-iso"), by = "word") |>
  group_by(CourseId, review_id) |>
  summarize(review_clean = paste(word, collapse = " "))
```

# Your turn

Explore the `en_coursera_reviews` dataset with the folks around you. Can you replicate my analyses, and try a few of your own? Here are some ideas: 

- Explore another subject (besides `data-science`)
- `Label` contains the numeric rating for the course. How does the review text differ between highly rated and low-rated courses?
- `stopwords::stopwords_getsources()` gives the available stopword dictionaries. How do these differ, and how do they change your results?
