---
title: "16-functions"
---

```{r}
library(tidyverse)
library(palmerpenguins) # load penguins data
```

# Load the Data

```{r}
unscaled_cancer <- read_csv("https://raw.githubusercontent.com/UBC-DSCI/introduction-to-datascience/refs/heads/main/data/wdbc_unscaled.csv")

unscaled_cancer
```

## Your turn: 3 functions

Turn the following code snippets into functions. Think about what each function does before you begin, and be sure to give each function an informative name.

1. `mean(is.na(x))`

1. `x / sum(x, na.rm = TRUE)`

1. `sd(x, na.rm = TRUE) / mean(x, na.rm = TRUE)`


## Your turn: `column_mean`

- Write a function called `column_mean` that takes a data set and column name (as a string) as inputs and returns the column mean as output. (*Hint:* access the column using `[[`)

- You should also include a `na.rm` argument and set the default to `TRUE` so that `NA`s are removed from the calculation by default.

- Test your function on the `mtcars` data set.

```
> column_mean(mtcars, "cyl")
[1] 6.1875
```

## Your turn: `scatterplot`

Write a plotting function that makes a scatterplot of any two quantitative variables, coloring the points by a 3rd categorical variable. 

Test your function with the following examples:

`scatterplot(unscaled_cancer, Radius, Texture, Class)`

`scatterplot(penguins, bill_length_mm, bill_depth_mm, species)`

## Your turn: `scatterplot` 2

Edit your `scatterplot` function to include an argument called `draw_line`. If `draw_line` is `TRUE`, your function should add a line of best fit to your scatterplot. Test your function with the following examples

`scatterplot(unscaled_cancer, Radius, Texture, Class, draw_line = FALSE)`

`scatterplot(penguins, bill_length_mm, bill_depth_mm, species, draw_line = TRUE)`

# More practice (if time)

## 1

Consider the following function and subsequent call of that function. What causes this error? How can you fix it?

```{r}
#| eval: false
summarize_species <- function(pattern = "Human") {
  starwars |>
    filter(species == pattern) |>
    summarize(
      num_people = n(),
      avg_height = mean(height, na.rm = TRUE)
    )
}

summarize_species(Wookiee)
```

## 2

You can tell R to print messages using `print("your message here")`. Edit the function above to first check if `pattern` is one of the `species` in the dataset. If it's not, your function should print an informative message and then `return()`. 

## 3 

Given a vector of birthdates, write a function to compute the age in years. See if your function works on the `easy` vector first, then try the `hard` vector. 

```{r}
hard <- c("15 Feb 1992", "10 March 1985", "03/28/2024", "09/30/2005")
easy <- mdy(c("2/15/1992", "3/10/1985", "3/28/2024", "9/30/2005"))
```
