---
title: "18: Iteration II"
---

```{r}
library(tidyverse)
```

# Load the Data

```{r}
unscaled_cancer <- read_csv("https://raw.githubusercontent.com/UBC-DSCI/introduction-to-datascience/refs/heads/main/data/wdbc_unscaled.csv")


bakeoff = read_csv("https://stat220-w26.github.io/data/bakeoff-episodes.csv") |>
  filter(series == 14)

# us_castaway_results
load(url("https://stat220-w26.github.io/data/combining-data-examples.Rda"))

# penguins
library(palmerpenguins)
```



## Try it: `map`

1. Edit the code chunk below so it returns a numeric vector (remove `eval: false`)

2. Edit the code chunk so it only `map`s to the numeric columns (3-12) of `unscaled cancer`

```{r}
#| eval: false
map(unscaled_cancer, mean)
```

3. `map` the `summary` function to all columns in the {palmerpenguins} `penguins` data

```{r}
# your code here
```

## Your turn: `penguins` and `map`

Using the `penguins` data, use `map` to calculate the `range` of a numeric variable and the `table` of a factor variable. (It may be helpful to first write a custom function for this output)

Your result should be a list (it will have length 8).

```{r}
# your code here
```

## More practice: survivor data

```{r}
us_castaway_results 
```

1. Write a function called `finalists` that takes the input of a survivor season (as a `numeric`) and outputs a string of the finalists' names for that season. The finalists' names should be separated with a comma. 

2. Use `map_chr` to return a character vector of finalists for seasons 31-40. 

## Your turn: `sample_finalists`

Some survivor seasons only had 16 players, while others had 22. This could result in some seasons being slightly under/overrepresented in our sample. Let's account for this.

Edit this experiment to instead randomly sample one player from each season (this results in 47 players) and then sample 20 players from the 47 random ones. 

(*Hint:* look at the `by` argument in `slice_sample`)

```{r}
# edit this function
sample_finalists = function(n){
    us_castaway_results %>%
    slice_sample(n = n) %>%
    filter(finalist) %>%
    count() %>%
    pull(n)
}
```

```{r}
# your iteration code here
```

# Permutation test


```{r}
unscaled_cancer %>%
  group_by(Class) %>%
  summarize(
    mean = mean(Radius)
  )
```

1. Define a function to run the experiment
    - Create a new column of `unscaled_cancer` called `class_shuffled`, which is a permutation of the original `Class` variable (*Hint:* see `sample` function)
    - Group by `class_shuffled` and compute the group means
    - Find the difference in the means
2. Repeat the experiment 1000 times, making sure to save the difference in means
3. Make a histogram of the simulated differences. How (un)likely is the difference we observed?


