---
title: "19-apis"
---

# TidyCensus

Install `tidycensus`

```{r}
#| eval: false
install.packages("tidycensus")
```

```{r}
library(tidycensus)
library(tidyverse)
```

## Example call:

```{r}
#| eval: false
acs_mn_2020 <- tidycensus::get_acs(
    year = 2020,
    state = "MN",
    geography = "county",
    variables = c("B01003_001", "B19013_001"),
    output = "wide",
    geometry = TRUE
)
```

## Storing your API key

1. Create a new text file in the same folder as your .rmd
2. Copy and paste your census key into the empty file
3. Save the file as `census_api_key.txt`

```{r}
#| eval: false
my_key <- readLines("census_api_key.txt")
```

and tell `tidycensus` what your API key is with:

```{r}
#| eval: false
census_api_key(my_key)
```

*Do not commit and push `census_api_key.txt` to github*

# httr2

Install if not already installed:

```{r}
#| eval: false

install.packages("httr2")
```

```{r}
library(httr2)
```

## Example call

```{r}
#| eval: false
hmong_state_request <- request("https://api.census.gov/data") %>% 
    req_url_path_append("2019") %>% 
    req_url_path_append("acs") %>% 
    req_url_path_append("acs1") %>% 
    req_url_query(get = c("NAME", "B02015_009E", "B02015_009M"), `for` = I("state:*"), key = census_api_key, .multi = "comma")
```

```{r}
#| eval: false
hmong_state_response <- req_perform(hmong_state_request)
```

```{r}
#| eval: false
hmong_state_tbl <- hmong_state_response %>%
  resp_body_json(simplifyVector = TRUE) %>%
  janitor::row_to_names(1) %>%
  as_tibble()

hmong_state_tbl
```


## Your Turn



- Edit the `httr` code to access a new variable of your choice
- Make a `httr` request to access the 1-year ACS data from 2018, 2019, 2021, and 2022. Make sure to save your results from each call! 
- Combine all years into a single dataset
- Make a time series plot with your chosen variable on the `y-axis`, year on the `x-axis`, colored by `state.` 
    - You may want to first filter to only a few states
    - You will need to do some cleaning of the data
    

