---
title: "21-scraping"
---

```{r}
#library(tidyverse)
#library(rvest)
```

# `paths_allowed`

```{r warning=FALSE}
#library(robotstxt)
#paths_allowed("http://www.zillow.com")
```

# Box Office Mojo

```{r}
#| eval: false 
#page <- read_html("https://www.boxofficemojo.com/year/2024/")

```

# Carleton Class Schedule 

<https://www.carleton.edu/catalog/current/search/?subject=STAT&term=25WI>

View the page source to try to find the html elements where this data is located (e.g. 'h1', 'p', 'table')

- Course number
- Course title
- Course description
- Course meetings
- Faculty
- Course meetings

```{r}
#listings = read_html("https://www.carleton.edu/catalog/current/search/?subject=STAT&term=25WI")
```

# SelectorGadget

- Use the SelectorGadget to explore http://www.imdb.com/chart/top

- What should the columns of our target dataset be? Do they correspond to any specific css selectors?

# Scraping IMDb Movie Page 

```{r}
#imdb <- read_html("http://www.imdb.com/chart/top")
```

```{r, eval=FALSE}
#| eval: false

#_______ <- imdb %>%
#  html_elements(".with-margin .ipc-title__text") %>%
#  html_text()

#_______ <- imdb %>%
#  html_elements(".cli-title-metadata-item:nth-child(1)") %>%
#  html_text()

#_______ <- imdb %>%
#  html_elements(".cli-title-metadata-item:nth-child(2)") %>%
#  html_text()

#_______ <- imdb %>%
#  html_elements(".cli-title-metadata-item:nth-child(3)") %>%
#  html_text()

#imdb_top_250 <- tibble(
  
#  )

#imdb_top_250
```

# Your Turn: IMDb TV Shows Page

In an R script:

  - Scrape the names, scores, and years of most popular TV shows on IMDB:
  [www.imdb.com/chart/tvmeter](http://www.imdb.com/chart/tvmeter)
  
  - Create a data frame called `tvshows` with the variables:
    `rank`, `title`, `stars`, `year`, `episodes`, `n_ratings`
    
  - Wrangle your resulting data so that all variable types are imported correctly
  
  - Use `write_csv` to save your file. If time, read it into the `21-scraping.rmd` and make a graph
  
