---
title: "25-shiny-design"
---

```{r}
library(tidyverse)
```

# Warm Up

Open the following shinyapps and explore a little bit. 

+ [Data Wrangling](https://psu-eberly.shinyapps.io/data_wrangling/)
+ [Biodiversity in NPS](https://abenedetti.shinyapps.io/bioNPS/)
+ [Hex Memory Game](https://dreamrs.shinyapps.io/memory-hex/)
+ [Lake Profile Dashboard](https://gallery.shinyapps.io/lake-profile-dashboard/)
+ [Classification Trees](https://asluby.shinyapps.io/GSS_classification/)

With the folks around you, discuss: 

1. Did any apps stand out as enjoyable or useful? 
2. What features or design choices make an app stand out? 
3. Are there any small tweaks you could make to make them more usable?

# AskAManager App 

Use the version you've been working on, or use a new starter app at [https://stat220-w26.github.io/files/25-starter-app.R](https://stat220-w26.github.io/files/25-starter-app.R) (it is also in the solutions repo)

## Part 1

- Move the initial data preparation code to an R **script** called `data-prep.R`
- Write the cleaned dataset to a CSV or .Rds file
    + `write_rds(manager_survey, "manager-survey/data/manager-survey.rds")`
- In the `app.R` file, load the data with `read_csv`
    + `manager_survey <- read_rds("data/manager-survey.rds")`

## Part 2

- Add a `sliderInput` to the appropriate tab panel

```{r}
#| eval: false
sliderInput(
  inputId = "ylim",
  label = _______ ,
  min = 0,
  value = c(0, 1000000),
  max = __________,
  width = "100%"
)
```

- Edit your ggplot code for the individual salary plot to display only the data points that are within these slider values

## Part 3

- Add a custom theme to the AskAManager app
- Update the ggplot themes to match

## (If Time)

Set up shinyapps.io account and deploy either (a) Portfolio 4 or (b) the AskAManager app


- You can deploy shiny apps online
    + using Posit's cloud server (free/fee) - https://www.shinyapps.io/
        - ["Getting started"](https://docs.posit.co/shinyapps.io/guide/getting_started/) guide will walk you through connecting RStudio to your shinyapps.io account
        - note: there is also now an option to deploy via Connect Cloud from your github. Feel free to try it!
    + creating a shiny server

- Your app and files should be in their own **folder** and the entire folder is what you "deploy". You do *not* want multiple `app.R` or `.Rmd` files in that folder! 


::: aside
Credit: [Mine Çetinkaya-Rundel's](https://vizdata.org/) Ask a Manager application exercise
:::
