---
title: "26-sql"
---

```{r}
#library(tidyverse)
#library(mdsr) # for connecting to database
#library(DBI) # database tools
```

# Warm Up 

Using `nycflights23::flights`, answer the following using {dplyr} code: 

- How many flights are in this dataset? 
- How many flights flew to MSP from JFK in 2023?
- What was the average flight delay for MSP flights in Sept 2023?

```{r}
#nycflights23::flights
```

# Intro to SQL

## Your turn: Set up database connection

```{r}
#| eval: false
#db <- dbConnect_scidb(dbname = "imdb")
```

```{r}
#| eval: false
#db %>%
#  dbGetQuery("SELECT * FROM kind_type
#             LIMIT 0,10;")
```

What `kind_type` corresponds to movies?


## Your turn: simple queries

1.  Find your favorite movie in the `title` table
2. Find [Viola Davis](https://www.imdb.com/name/nm0205626/?ref_=nv_sr_1) in the `name` table 
3.  Make note of Viola Davis's `person_id` 

## Your turn: finding data info

List the fields in the `name`, `cast_info`, and `char_name` tables. Do they have fields in common? How do you think these tables fit together?

```{r}
#| eval: false
#dbListTables(db)
#dbListFields(db, "title")
```

## Your turn: `JOIN`ing Viola Davis

1. Find all the rows in `cast_info` that correspond to Viola Davis as an actress

2. Add the names of the characters she played by `JOIN`ing the `char_name` table

## Your turn: full filmography

Find Viola Davis’s full filmography, in chronological order. Include each movie’s `title`, `production_year`, and the `name` of the character that she played.
