---
title: "Portfolio Project 1"
subtitle: "Posing and answering research questions"
---

```{r}
library(tidyverse)
library(tidytuesdayR)
```

# Overview

In your first portfolio project, you'll explore a dataset and pose your own research question.  You'll then apply what you've learned about `ggplot2` and design principles to answer the question.

Your final report should include

(1) A paragraph introduction to your data
(2) A section discussing your research question that includes:
    - An explicit statement of your question
    - Why that question is interesting/important
    - What you expect your answer to be to this question and why (this should be argued BEFORE digging deeply into the data; it's okay if you find out you were wrong!)
    - A graph that answers the question (or multiple, if more than one best answers the question)
    - A written answer to the question based on the graph(s), including whether your expectation was proven right or wrong
(3) A section describing your design choices, including
    - A paragraph telling me why you chose the geometries, aesthetics, scales, themes, etc. that you did
    - A brief reflection on whether it is possible other choices may have been valid as well, and whether changes there may have impacted your conclusions. You do NOT need to create other visuals to answer this part, but instead merely reflect in words.
    
Your question should involve at least two variables to answer. You should phrase the question in a way that is within the scope of inference of your data. For example, if you have an observational dataset, you shouldn’t phrase your question in a causal way. **Your questions shouldn't be identical to the example questions in the TidyTuesday repo pages.**

Your rendered file should *not* `echo` the code used to create the graphs, but the code in your `.qmd` should be clean and readable. Your report should not be long! If rendering to a PDF, I expect the entire report to be maybe 2-3 pages. 

# Options

## 1. World Languages

```{r}
endangered_status <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-12-23/endangered_status.csv")
families <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-12-23/families.csv")
languages <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-12-23/languages.csv")
# Download raw data and filter to endangered status
endangered_status <- 
  readr::read_csv("https://raw.githubusercontent.com/glottolog/glottolog-cldf/refs/heads/master/cldf/values.csv") |> 
  dplyr::filter(Parameter_ID == "aes") |> 
  dplyr::select(Language_ID, Value, Code_ID) |> 
  dplyr::rename(id = Language_ID,
                status_code = Value,
                status_label = Code_ID) |> 
  dplyr::mutate(status_label = stringr::str_replace(stringr::str_remove(status_label, "^aes-"), "_", " "))

# Download language and family data
fam_lgs <- 
  readr::read_csv("https://raw.githubusercontent.com/glottolog/glottolog-cldf/refs/heads/master/cldf/languages.csv")

# Filter and clean language family data
families <- 
  fam_lgs |> 
  dplyr::filter(Level == "family") |> 
  dplyr::select(ID, Name) |> 
  dplyr::rename(Family = Name) |> 
  dplyr::rename_with(stringr::str_to_lower, dplyr::everything())

# Filter and clean language data
languages <- 
  fam_lgs |> 
  dplyr::filter(Level == "language") |> 
  dplyr::select(ID, Name, Macroarea, Latitude, Longitude, ISO639P3code, Countries, Is_Isolate, Family_ID) |> 
  dplyr::rename_with(stringr::str_to_lower, dplyr::everything())

# Merge datasets together
languages <- merge(languages, endangered_status, by="id")
languages <- merge(languages, families, by.x="family_id", by.y="id")

glimpse(languages)
```

For more information about the variables in the dataset, see the [TidyTuesday repo page](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-12-23/readme.md).

## 2. Olympics

```{r}
olympics <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2026/2026-02-10/schedule.csv')
olympics
```

For more information about the variables in the dataset, see the [TidyTuesday repo page](https://github.com/rfordatascience/tidytuesday/tree/main/data/2026/2026-02-10).


## 3. NSF grant terminations

```{r}
nsf_terminations <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-05-06/nsf_terminations.csv')
nsf_terminations
```

For more information about the variables in the dataset, see the [TidyTuesday repo page](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-05-06/readme.md).

## 4. US Judges

```{r}
judges_appointments <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-10/judges_appointments.csv')
judges_people <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2025/2025-06-10/judges_people.csv')
judges <- merge(judges_appointments, judges_people, by="judge_id")

glimpse(judges)
```

For more information about the variables in the dataset, see the [TidyTuesday repo page](https://github.com/rfordatascience/tidytuesday/blob/main/data/2025/2025-06-10/readme.md).

## 5. Dungeons and Dragons Monsters

```{r}
dnd <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2025/2025-05-27/monsters.csv')

dnd
```

For more information about the variables in the dataset, see the [TidyTuesday repo page](https://github.com/rfordatascience/tidytuesday/tree/main/data/2025/2025-05-27).

## 6. BYO 

If none of the options listed here speak to you, you can also choose your own dataset. The data should have at least two quantitative and two categorical variables, at least eight variables total, and be "complicated" enough to demonstrate an advanced understanding of `ggplot` and design principles from class. I'd encourage you to run your idea by me before diving too deep! 

# Submission

Your submission should include at least one visualization, a written overview of your work, and all associated code. You should work in a `quarto` file and render to pdf. I will distribute GitHub skeleton repos that you should work within. You should commit and push your final code to GitHub. You'll need to submit a pdf to Gradescope as well.

# Rubric

A **successful** project will:

- [ ] Include all necessary components pushed to GitHub and uploaded to Gradescope
    - There should be at least two commits with substantial changes between the two
- [ ] Contain code, plot(s), and written summary in a single .qmd document
- [ ] The .qmd file contains all necessary code for recreation
- [ ] The .qmd file does not include unnecessary code
- [ ] The research question posed is appropriate, thoughtful, and multivariate (involve at least two variables)
- [ ] The graph(s) are appropriate to answer the research question
- [ ]  Meet minimum submission quality standards
    -   Very few grammatical mistakes, spelling mistakes, or typos
    -   Appropriate labels and font sizes
    -   A readable theme
    -   The rendered document does not contain any unnecessary content (package loading messages, warnings, etc.)

An **excellent** project will meet all of the requirements for a successful project, plus

- [ ] The question uses three variables from the data at once
- [ ] The description of the dataset uses inline R code
- [ ] The graph uses two or more geometry layers (this might be two layers of data, or a map + points, annotations, direct labeling with `geom_text`, etc.) OR includes a ggplot element that goes beyond what we discussed in class (eg., a new ggplot extension package or a geometry we didn't discuss or see in homework)
- [ ] Graph(s) use consistent themes that have been customized beyond a full theme (eg `theme_minimal` or `theme_bw`) 
- [ ] Graphs contain a custom `scale_` element (color, x/y axis, size, etc.)
- [ ] A clear history of commits that show your progress (e.g. commit 1: load data and make basic graph, commit 2: make graph #2, commit 3: change idea for graph #1, etc.)

This is an *individual* portfolio project. 
