---
title: "An Overview of Lego Sets on Rebrickable"
date: "Last updated on `r Sys.Date()`"
author: "Amanda Luby"
output: html_document
params:
  data: "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-06/sets.csv.gz"
  data_source: '[2022-09-09 repository](https://github.com/rfordatascience/tidytuesday/tree/master/data/2022/2022-09-06) on [Tidy Tuesday](https://github.com/rfordatascience/tidytuesday)'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

```{r}
#| echo: false
sets <- readr::read_csv(params$data)
```

# Overview

[Rebrickable](https://rebrickable.com) is a website that shows you which LEGO sets you can build from the LEGO sets and parts that you already own. To do this, Rebrickable maintains a database of the entire LEGO catalog. In this document, we’ll summarize the LEGO sets in this database.

The data set was originally obtained from the `r params$data_source`.

## Summarizing the Lego dataset

In the data set consists of `r nrow(sets)` LEGO sets (i.e., rows) and `r ncol(sets)` variables (i.e., columns). The data set includes LEGO sets from `r min(sets$year)` to `r max(sets$year)`. The average number of parts in a set was `r round(mean(sets$num_parts), 1)` with a standard deviation of `r round(sd(sets$num_parts), 2)`. However, there are `r sum(sets$num_parts == 0)` in the data set with 0 parts, making these summary statistics inaccurate.

Below is a scatterplot with smoother describing how the typical number of parts in a set has changed from `r min(sets$year)` to `r max(sets$year)`.

```{r}
#| warning: false


library(tidyverse)

ggplot(sets, aes(year, num_parts)) +
    geom_jitter(alpha = 0.2, size = .5) +
    geom_smooth(color = "skyblue") +
    scale_y_continuous(trans = "log10") +
    coord_fixed(ratio = 10) +
    labs(x = "Year", y = "Number of parts",
             title = "LEGO sets are getting larger over the years",
             caption = "Data source: brickable.com") +
    theme_light()
```

